---
title: "CropSyst Tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{cs_overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{tidyverse} 
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  fig.asp = 0.618,
  fig.width = 5,
  out.width = "70%",
  fig.retina = 3
)
```

```{r setup}
library(tidyverse)
library(cropsyst)
```

# Model construction

The constructor function `cs_setup()` is used to initialize an empty cropsyst model object to which additional model parameters can be added using the following functions:

- `cs_add_weather()` 
- `cs_add_soil()`
- `cs_add_canopy_cover()`
- `cs_add_crop()`
- `cs_add_irrigation()` (optional)

# Input Data and parameters

## Weather

Daily weather data is required and shouild be inlcuded as a data frame and follow the naming format below.

### Input Data Set

Input weather data should be at a continuous daily time scale and contain the following columns

- **date** Date of observation should ba a Date object
- **precip** Total daily precipitation (mm)
- **tmax** Maximum daily temperature (°C)
- **tmin** Minimum daily temperature (°C)
- **s_rad** Total daily solar radiation (MJ/m2)
- **rh_max** Maximum relative humidity (%)
- **rh_min** Minimum relative humidity (%)
- **wind_speed** Average daily wind speed (m/s)

An example dataset `cs_weather` is provided with cropsyst and is shown below:
```{r weather_data, echo=FALSE}
head(cs_weather)
```

### Additional Parameters

- **latitude** location latitude (degrees)
- **elevation** location elevation (m)
- **screening_height** wind speed measurement height (m)

## Soil

### Input Data Set

Soil data should contain the following columns

- **horizon_thickness** Thickness of soil horizon (m)
- **sand_pct** Percent of horizon composed of sand (%)
- **clay_pct** Percent of horizon composed of clay (%)
- **silt_pct** Percent of horizon composed of silt (%)
- **initial_wc** Initial water content on first day of planting (m3/m3)


An example dataset `cs_soil` is provided with cropsyst and is shown below:
```{r soil_data, echo=FALSE}
head(cs_soil)
```

### Additional Parameters

- **thickness_evaporative_layer** Thickness of soil evaporative layer (m)

## Canopy Cover

### Input Data Set

- **gcc** Green canopy cover
- **tcc** Total canopy cover
- **dap** Day after planting

An example dataset `cs_canopy_cover` is provided with cropsyst and is shown below:
```{r canopy_cover, echo=FALSE}
head(cs_canopy_cover)
```

## Irrigation

Cropsyst simulations are able to use 3 irrigation scenarios

1. Scheduled irrigation
2. Auto irrigation
3. No irrigation

### Input Data Set
- **dap** Day after planting
- **irrigation** Irrigation (mm)

An example dataset `cs_irrigation` is provided with cropsyst and is shown below:
```{r irrigation, echo=FALSE}
head(cs_irrigation)
```

### Additional Parameters

- **irrigated_fraction** fraction of land irrigated in simulated area
- **max_allowable_depletion** Max soil water dfepletion which triggers irrigation
- **min_leaf_water_potential** Min leaf water potential which triggers irrigation
- **max_system_capacity** Max irrigation daily capacity
- **dae_first_irrigation** Day after emergence of first irrigation

## Crop Parameters

- **planting_date** Planting date
- **harvest_date** Harvest date
- **midseason_et_crop_coef** mid season ET crop coefficient
- **tree_fruit** logical TRUE if tree fruit crop
- **fruit_harvest_date** harvest date of tree fruit
- **max_crop_height** max crop height (m)
- **max_root_depth** max root depth (m)
- **perrenial** logical TRUE if perennial
- **max_water_uptake** max water uptake
- **transpiration_use_eff** transpiration use efficiency
- **C3** C3 = TRUE C4 = FALSE


# Model Setup

```{r cropsyst_object}
cs_input <- cs_setup() %>% 
  cs_add_weather(weather = cs_weather, 
                 latitude = 35.2, 
                 elevation = 1170,
                 screening_height = 2) %>% 
  cs_add_soil(soil = cs_soil, 
              thickness_evaporative_layer = 0.005) %>% 
  cs_add_canopy_cover(cs_canopy_cover) %>% 
  cs_add_irrigation(cs_irrigation, 
                    irrigated_fraction = 0.0001) %>% 
  cs_add_crop(planting_date = ydoy(2016, 131), 
              season_end_date = ydoy(2016, 131+156-1),
              midseason_et_crop_coef = 1.15,
              tree_fruit = FALSE,
              fruit_harvest_date = ydoy(2016, 131+153),
              max_crop_height = 2,
              max_root_depth = 2, 
              perrenial = FALSE, 
              max_water_uptake = 14, 
              transpiration_use_eff = 7, 
              C3 = FALSE)
```

# Run Model

```{r run_model}

cs_out <- cs_input %>% cs_run()

```

# Model Output

The model output maintains the form of the model input with weather, irr_soil, and non_irr_soil as data frames and params as a list of model input parameters.

```{r model_output_names}
names(cs_out)
```


## Potential ET

```{r potential_et}

cs_out$weather %>% 
  ggplot() +
  geom_line(aes(x = dap, y = potential_crop_et))


```

## Crop ET Coeficcient

```{r crop_coef}

cs_out$weather %>% 
  ggplot() +
  geom_line(aes(x = dap, y = et_crop_coef)) 


```

## Crop Growth

```{r crop_growth}

cs_out$weather %>% 
  ggplot() +
  geom_line(aes(x = dap, y = -root_depth, color = "Root Depth")) +
  geom_line(aes(x = dap, y = crop_height, color = "Crop Height")) +
  labs(x = "Days after planting", y = "Height (m)")


```

## Biomass

```{r biomass_production}

cs_out$weather %>% 
  ggplot() +
  geom_line(aes(x = dap, y = potential_biomass_prod, color = "potential")) +
  geom_line(aes(x = dap, y = biomass_production, color = "actual"))

cs_out$weather %>% 
  ggplot() +
  geom_line(aes(x = dap, y = cum_biomass_production))


```

## Transpiration

```{r transpiration}


cs_out$weather %>% 
  ggplot() +
  geom_line(aes(x = dap, y = attainable_transpiration, color = "Attainable")) +
  geom_line(aes(x = dap, y = potential_transpiration, color = "Potential")) +
  geom_line(aes(x = dap, y = actual_transpiration, color = "Actual"))


```

## Soil Water Evaporation

```{r soil_water_evap}


cs_out$weather %>% 
  ggplot() +
  geom_line(aes(x = dap, y = attainable_soil_water_evaporation, color = "Attainable")) +
  geom_line(aes(x = dap, y = irr_zone_soil_water_evap, color = "Irrigated")) +
  geom_line(aes(x = dap, y = non_irr_zone_soil_water_evap, color = "Non-Irrigated"))


```

# Leaf Water Potential

```{r leaf_water_pot}


cs_out$weather %>% 
  ggplot() +
  geom_line(aes(x = dap, y = leaf_water_potential)) 


```

# Actual Evapotranspiration

```{r actual_et}
cs_out$weather %>% 
  ggplot() +
  geom_line(aes(x = dap, y = actual_evapotranspiration)) 
```
s
